Towards Spherical Robots for Mobile Mapping in Human Made Environments\tnoteref{mytitlenote}

### 2021-08-20 Latex/sections/introduction.tex
added comma, look of pushing works
Incorporated both reviewers comments

Today's robots for mobile mapping come in all shapes and sizes.
State of the art for urban environments are laser scanners mounted to cars.
Smaller robotic systems are particularly used when cars no longer have access.
Examples for this are human operated systems such as Zebedee~\cite{Bosse2012-zebedee}, a small Hokuyo 2D scanner on a spring, that is carried through the environment, VILMA~\cite{JPRS2016}, a rolling FARO scanner operating in profiler mode, RADLER~\cite{Borrmann2020-RADLER}, a SICK 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack} or~\cite{WWWLeicaBackpack}.
Recently more and more autonomous systems gained maturity.
A stunning example is Boston Dynamics' quadruped ``Spot'' that autonomously navigates and maps human environments~\cite{SpotRobot}.
Also, the mobile mapping approaches implemented on the ANYmal platform such as~\cite{Fankhauser2018-ANYmal} were very successful.
Of all these formats, one has not been explored thoroughly in the scientific community: The spherical mobile mapping robot.
Yet this provides some very promising advantages over the other formats.
For one, the locomotion of a spherical robot inherently results in rotation.
That way, a sensor fixed inside the spherical structure will cover the entire environment, given the required locomotion without the need for additional actuators for the sensors.
This requires a solution for the spherical simultaneous localization and mapping (SLAM) problem given the six degrees of freedom of the robot.
Secondly, a spherical shell that encloses all sensors protects these from possible hazardous environments.  
For example, the shell stops any dust that deteriorates sensors or actuators when settling at sensitive locations.
In contrast to an usual enclosing the shell can separate the sensors entirely from the environment without the need for a number of points-of-connection. 
A strict requirement then is that the shell is very durable.
This is particularly useful for unknown or dangerous environments.
E.g., old buildings that are in danger of collapsing, narrow underground funnels, construction sites, or mining shafts. 
The spherical format is, in fact, also suited for space applications.
In the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, such a robot is proposed that is to be lowered into a lunar cave and create a 3D map of the environment. The authors choose this format as the moon regolith is known to damage instruments and other components.
They also present an approach to protect the shell from accumulating dust and dirt.
However, the spherical format also comes with some disadvantages.
Lidar sensors often have a minimum scan distance, resulting in a less dense (or empty) point cloud when the scanner looks on the ground, whereas density is higher when looking in other directions.
The ground itself is likely to be less populated with points, due to weak angles of incidence while mapping it.
Furthermore, relying on IMU based odometry as a localization technique alone yields inaccurate and noisy pose measurements.
Therefore, a robust registration procedure is needed for spherical robots, that is able to cope with vast differences is point cloud density and high noise regarding pose measurements.

>>>>

Today's robots for mobile mapping come in all shapes and sizes.
State of the art for urban environments are laser scanners mounted to cars.
Smaller robotic systems are particularly used when cars no longer have access.
Examples for this are human operated systems such as Zebedee~\cite{Bosse2012-zebedee}, a small Hokuyo 2D scanner on a spring, that is carried through the environment, VILMA~\cite{JPRS2016}, a rolling FARO scanner operating in profiler mode, RADLER~\cite{Borrmann2020-RADLER}, a SICK 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack} or~\cite{WWWLeicaBackpack}.
Recently more and more autonomous systems gained maturity.
A stunning example is Boston Dynamics' quadruped ``Spot'' that autonomously navigates and maps human environments~\cite{SpotRobot}.
Also, the mobile mapping approaches implemented on the ANYmal platform such as~\cite{Fankhauser2018-ANYmal} were very successful.
Of all these formats, one has not been explored thoroughly in the scientific community: The spherical mobile mapping robot.
Yet this provides some very promising advantages over the other formats.
For one, the locomotion of a spherical robot inherently results in rotation.
That way, a sensor fixed inside the spherical structure will cover the entire environment, given the required locomotion without the need for additional actuators for the sensors.
This requires a solution for the spherical simultaneous localization and mapping (SLAM) problem, given the six degrees of freedom of the robot.
Secondly, a spherical shell that encloses all sensors protects these from possible hazardous environments.  
For example, the shell stops any dust that deteriorates sensors or actuators when settling at sensitive locations.
In contrast to an usual enclosing the shell can separate the sensors entirely from the environment without the need for a number of points-of-connection. 
A strict requirement then is that the shell is very durable.
This is particularly useful for unknown or dangerous environments.
E.g., old buildings that are in danger of collapsing, narrow underground tunnels, construction sites, or mining shafts. 
The spherical format is, in fact, also suited for space applications.
In the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, such a robot is proposed that is to be lowered into a lunar cave and create a 3D map of the environment. The authors choose this format as the moon regolith is known to damage instruments and other components.
They also present an approach to protect the shell from accumulating dust and dirt.
However, the spherical format also comes with some disadvantages.
Lidar sensors often have a minimum scan distance, resulting in a less dense (or empty) point cloud when the scanner looks on the ground, whereas density is higher when looking in other directions.
The ground itself is likely to be less populated with points, due to weak angles of incidence while mapping it.
Furthermore, relying on IMU based odometry as a localization technique alone yields inaccurate and noisy pose measurements.
Therefore, a robust registration procedure is needed for spherical robots, that is able to cope with vast differences is point cloud density and high noise regarding pose measurements.

### 2021-08-20 Latex/sections/realData.tex
images moved
Incorporated both reviewers comments

The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor the simulation uses.
We put the laser scanner into an acrylic glass spherical shell together with three IMUs  (same specifications as in the first dataset) and manually roll it with a slow, constant motion in a home indoor environment.
Thus, this time nearly all six degrees of freedom are used for optimization.
We exclude a translation in the global up- or downwards direction as the sphere rolls on the ground surface, without possibility for up- or downwards motion-
Figure~\ref{fig:prototype} shows the prototype.
The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the transformation defined by the recorded poses.
Although the robot has been moved carefully, you see that pose error accumulates in such a way that the same wall is sensed multiple times.

\begin{figure}
  \includegraphics[width=0.49\linewidth]{./images/photoReal1.jpeg}
  \includegraphics[width=0.50\linewidth]{./images/photoReal3.jpeg}
  \caption{Prototype used for data acquisition with the rolling sphere. The main payload is the Livox Mid-100 laser scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and the breakout box of the laser scanner. }
  \label{fig:prototype}
\end{figure}

>>>>

The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor the simulation uses.
We put the laser scanner into an acrylic glass spherical shell together with three IMUs  (same specifications as in the first dataset) and manually roll it with a slow, constant motion in a home indoor environment.
Thus, this time nearly all six degrees of freedom are used for optimization.
We exclude a translation in the global up- or downwards direction as the sphere rolls on the ground surface, without possibility for up- or downwards motion-
figure~\ref{fig:prototype} shows the prototype.
The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the transformation defined by the recorded poses.
Although the robot has been moved carefully, you see that pose error accumulates in such a way that the same wall is sensed multiple times.

\begin{figure}
  \includegraphics[width=0.495\linewidth]{./images/photoReal1.jpeg}\hfill
  \includegraphics[width=0.495\linewidth]{./images/photoReal3.jpeg}\\
  \includegraphics[width=0.495\linewidth]{./images/ProtoRoll1.png}\hfill
  \includegraphics[width=0.495\linewidth]{./images/ProtoRoll2.png}
  \caption{Top: Prototype used for data acquisition with the rolling sphere. The main payload is the Livox Mid-100 laser scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and the breakout box of the laser scanner. Below: Rendering of the simulated robot, a Livox laser scanner insider a spherical shell. }
  \label{fig:prototype}
\end{figure}

### 2021-08-20 Latex/sections/registration.tex
Incorporated both reviewers comments

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{images/project}
	\caption{Illustration of the Hesse- and polygon projection distance. The point p (red) gets projected onto the infinitely extending global plane. Both the point projection (green) and the global planes convex hull (blue) get projected into 2D space (grey). The Hesse distance, i.e. the shortest distance to the infinitely extending plane (red), is shown as well as the minimum distance from the 2D point projection to the polygon (purple).}
	\label{fig:proj}
\end{figure*}

@@

The simplicity of the Hesse distance is at contrast with its inability to take the extent of the plane into account.
Therefore, a second distance model is introduced: the polygon projection distance (PPD).
The convex hull represents the extend of a plane by forming a polygon with the outer most points that are assigned to a given plane. 
It is able to represent the expansion of the plane in all directions, and thus is utilized as a distance model.
To find the PPD, first a corresponding point gets projected onto the infinitely extending plane representation given by the Hesse form, i.e. $\widetilde{T}(\vec{p}_k)$ is found from eq.~\eqref{eq:projection} (green point in~\ref{fig:proj}).
Then, the 3D polygon, i.e. the points that make up the convex hull, as well as the corresponding 3D point projection, are projected again into a 2D vector space, using the maximal component of the normal vector of the plane.
E.g. in figure~\ref{fig:proj}, the blue planes normal vector has its mayor component in z-direction (upwards), thus the projection polygon distance is calculated in the xy-plane.
Using the maximum of the absolute magnitude of the normal vector ensures that the most sensible 2D projection is used for every direction.
Sunday et al.~\cite{pga} presents an optimal algorithm to check whether the point projection lies inside the polygon in 2D. 
If the point is inside the polygon, its PPD is set to zero.
If, however, the point is outside the polygon, the shortest distance to the polygon is calculated by looking at the minimum distance of the $i$-th point to each line segment, making up the convex hull of the $k$-th plane in 3D.
Let the polygon $P_k$ be made up of line segments $s_{m, j}$.
The line segment $s_{m, j}$ consists of the points $p_m$ and $p_j$.
The line segment is parameterized by
\begin{align}
\label{eq:line}
	s_{m, j}(t) = p_m + t \cdot \left( p_j - p_m \right) ,  
\end{align}
where $t \in \left[ 0, 1 \right]$. 
We set up a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
\begin{align}
	\label{eq:dist}
	d_{m, j}(t) = \lVert s_{m, j}(t) - p_i \rVert 
\end{align} 
It is now possible to find the shortest distance to the line segment by finding the argument of the function that minimizes this distance.
Therefore, we set
\begin{align}
	\pardiff{t} d_{m, j}(t)^2 \overset{!}{=} 0 ,
\end{align}  
resulting in 
\begin{align}
	t_0 = \frac{\left( p_j - p_m \right) \cdot \left( p_i - p_m \right)}{\left( p_j - p_m \right)^2} .
\end{align}
Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
In that case, the projection of the point $p_i$ onto the line given by eq.~\ref{eq:line} is not between $p_j$ and $p_m$. 
Instead, its projection onto the line falls outside of the segment.
By constraining $t_0$ we get 
\begin{align}
\label{eq:minmax}
	\hat t = \min \left( \max \left( t_0, 0\right) , 1 \right) ,
\end{align}
which is the argument that gives the shortest distance to the line segment, when inserted to eq. \ref{eq:dist}.
We find the PPD from the point $p_i$ to the polygon $P_k$ by calculating the minimum shortest distance over all line segments that make up the polygon.

@@

The presented algorithm employs LPC on each individual line scan.
The idea is to find planar features utilizing an optimized approximate k nearest neighbor (AKNN) search for normal calculation as in~\cite{aknnbbs}.
Thus, normals are calculated for every point in the respective line scan.
Then, the local normal model is combined with a region growing approach.
The region growing is implemented by storing the k-nearest neighbors in a queue, which represents the iteration order of the clustering algorithm.
A region grows if one of these neighbors has a distance shorter then a threshold $d_{growth}$ and additionally has a similar normal.
We define similarity of two normals via the angle between them, which is:
\begin{align}
\label{eq:smallestalpha}
	\alpha(\vec{n}_1, \vec{n}_2) = \arccos(\vec{n}_1 \cdot \vec{n}_2) \; .
\end{align}
However, the \textit{smallest} angle between two \textit{normal} vectors is 
\begin{align}
\label{eq:alphahat}
\hat \alpha(\vec{n}_1, \vec{n}_2) = 
\begin{cases}
	2\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{3}{2} \pi \\
	\alpha(\vec{n}_1, \vec{n}_2) - \pi & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \pi \\
	\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{\pi}{2} 
\end{cases}
\; ,
\end{align}
because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
We say that two normals are similar if the angle $\hat \alpha$ from eq. \ref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{images/clustering}
	\caption{(Left) Point normals using the AKNN method with $K = 20$. (Middle) Result of the region growing, before applying the second filter. (Right) Resulting clusters after applying the second filter, using $N_{c,min} = 300$. }
	\label{fig:clustering}
\end{figure*}

Point-to-Plane correspondences are now established.
The distance models (Hesse and PPD) are utilized to check whether a cluster overlaps with any of the global planes, i.e. any point in the cluster has a Hesse distance smaller than some threshold $\epsilon_H$ and a PPD smaller than some additional threshold $\epsilon_P$, to the global plane.
We observe the noise level around the expected planes by the naked eye. 
Then we set the thresholds accordingly. 
For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \ref{eq:alphahat}.
Finally, a correspondence is established between all points in the cluster and the global plane, if the minimum angle between their normals is smaller than $\epsilon_\alpha$.

>>>>

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{images/project}
	\caption{Illustration of the Hesse- and polygon projection distance. The point p (red) gets projected onto the infinitely extending global plane. Both the point projection (green) and the global planes convex hull (blue) get projected into 2D space (grey). The Hesse distance, i.e. the shortest distance to the infinitely extending plane (red), is shown as well as the minimum distance from the 3D point projection to the polygon (purple).}
	\label{fig:proj}
\end{figure*}

@@

The simplicity of the Hesse distance is at contrast with its inability to take the extent of the plane into account.
Therefore, a second distance model is introduced: the polygon projection distance (PPD).
The convex hull represents the extend of a plane by forming a polygon with the outer most points that are assigned to a given plane. 
It is able to represent the expansion of the plane in all directions, and thus is utilized as a distance model.
To find the PPD, first a corresponding point gets projected onto the infinitely extending plane representation given by the Hesse form, i.e. $\widetilde{T}(\vec{p}_k)$ is found from eq.~\eqref{eq:projection} (green point in figure~\ref{fig:proj}).
Then, the 3D polygon, i.e. the points that make up the convex hull, as well as the corresponding 3D point projection, are projected again into a 2D vector space, using the maximal component of the normal vector of the plane.
E.g. in figure~\ref{fig:proj}, the blue planes normal vector has its mayor component in z-direction (upwards), thus the projection polygon distance is calculated in the xy-plane.
Using the maximum of the absolute magnitude of the normal vector ensures that the most sensible 2D projection is used for every direction.
Sunday et al.~\cite{pga} presents an optimal algorithm to check whether the point projection lies inside the polygon in 2D. 
If the point is inside the polygon, its PPD is set to zero.
If, however, the point is outside the polygon, the shortest distance to the polygon is calculated by looking at the minimum distance of the $i$-th point to each line segment, making up the convex hull of the $k$-th plane in 3D.
Let the polygon $P_k$ be made up of line segments $s_{m, j}$.
The line segment $s_{m, j}$ consists of the points $p_m$ and $p_j$.
The line segment is parameterized by
\begin{align}
\label{eq:line}
	s_{m, j}(t) = p_m + t \cdot \left( p_j - p_m \right) ,  
\end{align}
where $t \in \left[ 0, 1 \right]$. 
We set up a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
\begin{align}
	\label{eq:dist}
	d_{m, j}(t) = \lVert s_{m, j}(t) - p_i \rVert 
\end{align} 
It is now possible to find the shortest distance to the line segment by finding the argument of the function that minimizes this distance.
Therefore, we set
\begin{align}
	\pardiff{t} d_{m, j}(t)^2 \overset{!}{=} 0 ,
\end{align}  
resulting in 
\begin{align}
	t_0 = \frac{\left( p_j - p_m \right) \cdot \left( p_i - p_m \right)}{\left( p_j - p_m \right)^2} .
\end{align}
Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
In that case, the projection of the point $p_i$ onto the line given by eq.~\eqref{eq:line} is not between $p_j$ and $p_m$. 
Instead, its projection onto the line falls outside of the segment.
By constraining $t_0$ we get 
\begin{align}
\label{eq:minmax}
	\hat t = \min \left( \max \left( t_0, 0\right) , 1 \right) ,
\end{align}
which is the argument that gives the shortest distance to the line segment, when inserted to eq. \eqref{eq:dist}.
We find the PPD from the point $p_i$ to the polygon $P_k$ by calculating the minimum shortest distance over all line segments that make up the polygon.

@@

The presented algorithm employs LPC on each individual line scan.
The idea is to find planar features utilizing an optimized approximate k nearest neighbor (AKNN) search for normal calculation as in~\cite{aknnbbs}.
Thus, normals are calculated for every point in the respective line scan.
Then, the local normal model is combined with a region growing approach.
The region growing is implemented by storing the k-nearest neighbors in a queue, which represents the iteration order of the clustering algorithm.
A region grows if one of these neighbors has a distance shorter then a threshold $d_{growth}$ and additionally has a similar normal.
We define similarity of two normals via the angle between them, which is:
\begin{align}
\label{eq:smallestalpha}
	\alpha(\vec{n}_1, \vec{n}_2) = \arccos(\vec{n}_1 \cdot \vec{n}_2) \; .
\end{align}
However, the \textit{smallest} angle between two \textit{normal} vectors is 
\begin{align}
\label{eq:alphahat}
\hat \alpha(\vec{n}_1, \vec{n}_2) = 
\begin{cases}
	2\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{3}{2} \pi \\
	\alpha(\vec{n}_1, \vec{n}_2) - \pi & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \pi \\
	\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{\pi}{2} 
\end{cases}
\; ,
\end{align}
because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
We say that two normals are similar if the angle $\hat \alpha$ from eq. \eqref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.

@@

\begin{figure*}
	\includegraphics[width=\linewidth]{images/clustering}
	\caption{(Left) Point normals using the AKNN method with $K = 20$ (number of nearest neighbors).
	(Middle) Result of the region growing, before applying the second filter.
	(Right) Resulting clusters after applying the second filter, using $N_{c,min} = 300$. 
	In each image, one color represents a segment.}
	\label{fig:clustering}
\end{figure*}

Point-to-Plane correspondences are now established.
The distance models (Hesse and PPD) are utilized to check whether a cluster overlaps with any of the global planes, i.e. any point in the cluster has a Hesse distance smaller than some threshold $\epsilon_H$ and a PPD smaller than some additional threshold $\epsilon_P$, to the global plane.
We observe the noise level around the expected planes by the naked eye. 
Then we set the thresholds accordingly. 
For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \eqref{eq:alphahat}.
Finally, a correspondence is established between all points in the cluster and the global plane, if the minimum angle between their normals is smaller than $\epsilon_\alpha$.

### 2021-08-20 Latex/sections/results.tex
yt link added to pdf
Incorporated both reviewers comments

The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. Figure~\ref{fig:simulatedDatasets}) without further processing.
Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.

@@

Figure~\ref{fig:jasperhome} shows the results obtained before and after employing the presented algorithm on the dataset, acquired by the prototype in figure~\ref{fig:prototype}.
We again combine twenty temporally successive scans into one meta-scan, using the scan at the median index as a reference coordinate system, which is then globally registered.
We used the first seven of these meta-scans, corresponding to the first full rotation of the spher, to extract global planes.
All subsequent meta-scans are then matched against the initial global plane model.
We use both optimizations from section \ref{ssec:furtheropts}, introducing a global up/down lock for gradient descent and applying sequential iteration to eliminate accumulating pose error.

\begin{figure*}
        \centering
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_corner}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_corner}\\
	\includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_color}\hfill
	\includegraphics[width=0.495\textwidth]{./images/jasper_corr_color}\\
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_top}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_top}
        \caption{The point cloud acquired by the rolling sphere before (left) and after (right) applying the plane based registration. View from the interior (top) and a birds-eye view (bottom). Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$. An animation of the registration process is given at \url{https://youtu.be/hYR3xLhIdsE}.}
        \label{fig:jasperhome}
\end{figure*}
After the registration, the walls of the environment are no longer ambiguous, since all the subsequent scans are matched with the initial plane representation.
When considering the resulting path, it looks distorted.
This is because for some scans, especially empty ones where the sensor points to the ground, no planar features got matched with the global plane model, thus their pose is not optimized.
Improving the path quality is an objective for future works.

>>>>

The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. figure~\ref{fig:simulatedDatasets}) without further processing.
Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.

@@

Figure~\ref{fig:jasperhome} shows the results obtained before and after employing the presented algorithm on the dataset, acquired by the prototype in figure~\ref{fig:prototype}.
We again combine twenty temporally successive scans into one meta-scan, using the scan at the median index as a reference coordinate system, which is then globally registered.
We used the first seven of these meta-scans, corresponding to the first full rotation of the sphere, to extract global planes.
All subsequent meta-scans are then matched against the initial global plane model.
We use both optimizations from section \ref{ssec:furtheropts}, introducing a global up/down lock for gradient descent and applying sequential iteration to eliminate accumulating pose error.

\begin{figure*}
        \centering
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_corner}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_corner}\\
	\includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_color}\hfill
	\includegraphics[width=0.495\textwidth]{./images/jasper_corr_color}\\
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_top}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_top}
        \caption{The point cloud acquired by the rolling sphere before (left) and after (right) applying the plane based registration. View from the interior (top) and a birds-eye view (bottom). Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$. An animation of the registration process is given at \url{https://youtu.be/2J8y7MN10JM}.}
        \label{fig:jasperhome}
\end{figure*}
After the registration, the walls of the environment are no longer ambiguous, since all the subsequent scans are matched with the initial plane representation.
When considering the resulting path, it looks distorted.
This is because for some scans, especially empty ones where the sensor points to the ground, no planar features got matched with the global plane model, thus their pose is not optimized.
Improving the path quality is an objective for future works.

### 2021-08-20 Latex/sections/stateOfTheArt.tex
Incorporated both reviewers comments

While there exist camera-based approaches~\cite{9233654}, to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before.
However, laser-based SLAM algorithms for motions in six degrees of freedom (DoF) have been thoroughly studied.
For outdoor environments~\cite{JFR2006} provides a first baseline.
Adding a heuristic for closed-loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
Zhang et al. propose a real-time solution to the SLAM problem in~\cite{Zhang2014}.
They achieve the performance at a lower computational load by dividing the SLAM algorithm into two different algorithms: one performing odometry at a high frequency but low fidelity and another running at a lower frequency performing fine matching and registration of the point clouds.
More recently Dröschel et al. also propose an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.
Another intriguing example is the NavVis VLX system~\cite{navvis}.
They perform real time SLAM on a mobile, wearable platform, producing colored point clouds with high accuracy by combining it with a camera system.
However, this platform has to be moved by a human operator. 
They also employ artificial markers that can be placed in the environment, which get recognized by the camera system to further improve registration accuracy.
Obviously, this is not possible in inaccessible, or dangerous environments.

@@

LOAM~\cite{Zhang2014} is the baseline-algorithm that provides a real-time and lowdrift solution based on two parallel registration algorithms using planes and lines.
Unfortunately, it is not open-source anymore.
LOAM livox extends the LOAM framework to the the rotating prism scanner with small FoV~\cite{Lin_2020}.
Zhou, Wang, and Kaess write that it also adopts the parallel computing to achieve real-time global registration.
Parallel computing needs a powerful CPU, it may be not suitable for an embedded system which has limited computational resources~\cite{Zhou_2021}.
Thus, they extend their smoothing and mapping to the LiDAR and planes case.
In their experiments they used a VLP-16 LiDAR to collect indoor datasets.

Further recent planar SLAM approaches include~\cite{Jung_2015,Grant_2018,Geneva_2018, Zhou_2021,   wei2021groundslam}.
While Wei et al. uses only the ground plane in outdoor experiments.
Indoors, Jung et al. used in 2015 several Hokuyo laser scanners for their kinematic scanning system~\cite{Jung_2015}, while Grant used a single Velodyne HDL-32E sensor with 32 rotating laser/receiver pairs mounted on a backpack and walked through different environments~\cite{Grant_2018}.
The LIPS system~\cite{Geneva_2018} employes a so-called Closest Point Plane Representation with an Anchor Plane Factor.
RANSAC is used to find the places.
Their system couples an eight channel Quanergy M8 LiDAR operating at 10Hz with a Microstrain 3DM-GX3-25 IMU attached to the bottom of the LiDAR operating at 500Hz.

>>>>

While there exist camera-based approaches~\cite{9233654}, to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before.
However, laser-based SLAM algorithms for motions in six degrees of freedom (DoF) have been thoroughly studied.
For outdoor environments~\cite{JFR2006} provides a first baseline.
Adding a heuristic for closed-loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
Thereby they reduce the position and orientation error of the scanning poses by orders of magnitude in the range of centimeters and hundreds of a degree.
Zhang et al. propose a real-time solution to the SLAM problem in~\cite{Zhang2014}.
They achieve the performance at a lower computational load by dividing the SLAM algorithm into two different algorithms: one performing odometry at a high frequency but low fidelity and another running at a lower frequency performing fine matching and registration of the point clouds.
More recently Dröschel et al. also propose an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.
Another intriguing example is the NavVis VLX system~\cite{navvis}.
They perform real time SLAM on a mobile, wearable platform, producing colored point clouds with high accuracy by combining it with a camera system.
However, this platform has to be moved by a human operator. 
In particular, they are able to achieve one sigma of measurements within three millimeters difference the measurements of a ground truth terrestrial laser scan~\cite{navvisAcc}. 
They also employ artificial markers that can be placed in the environment, which get recognized by the camera system to further improve registration accuracy.
Obviously, this is not possible in inaccessible, or dangerous environments.

@@

Lidar odometry and mapping (LOAM)~\cite{Zhang2014} is the baseline-algorithm that provides a real-time and lowdrift solution based on two parallel registration algorithms using planes and lines.
Unfortunately, it is not open-source anymore.
LOAM livox extends the LOAM framework to the rotating prism scanner with small FoV~\cite{Lin_2020}.
Zhou, Wang, and Kaess write that it also adopts the parallel computing to achieve real-time global registration.
Parallel computing needs a powerful CPU, it may be not suitable for an embedded system which has limited computational resources~\cite{Zhou_2021}.
Thus, they extend their smoothing and mapping to the LiDAR and planes case.
In their experiments they used a VLP-16 LiDAR to collect indoor datasets.

Further recent planar SLAM approaches include~\cite{Jung_2015,Grant_2018,Geneva_2018, Zhou_2021,   wei2021groundslam}.
While Wei et al. uses only the ground plane in outdoor experiments.
Indoors, Jung et al. used in 2015 several Hokuyo laser scanners for their kinematic scanning system~\cite{Jung_2015}, while Grant used a single Velodyne HDL-32E sensor with 32 rotating laser/receiver pairs mounted on a backpack and walked through different environments~\cite{Grant_2018}.
The LIPS system~\cite{Geneva_2018} employes a so-called Closest Point Plane Representation with an Anchor Plane Factor.
Random sample consensus (RANSAC) is used to find the planes.
Their system couples an eight channel Quanergy M8 LiDAR operating at 10Hz with a Microstrain 3DM-GX3-25 IMU attached to the bottom of the LiDAR operating at 500Hz.
